---
title: "R Dataframes"
output: html_notebook
---

# Lab \| Exploring the `sample_superstore` Dataset

In this lab, you will work with the `sample_superstore` dataset to practice creating, inspecting, and manipulating dataframes. Follow the steps below to complete the tasks.

## Step 1: Importing the Dataset

-   Load the `sample_superstore` dataset into a dataframe. You can find this dataset in the `datasets` package or download it as a CSV file from online sources.

-   Save the dataframe as `superstore`.

    **Hint:** Use the `read.csv()` or `read_excel()` function to import the dataset.
```{r}
df <- read.csv("dataset/Sample - Superstore.csv")
head(df)
```
    

## Step 2: Inspecting the Dataframe

-   Display the first 10 rows of the dataframe using the `head()` function.
```{r}
head(df, 10)
```

-   Use the `str()` function to inspect the structure of the dataframe. What are the data types of the columns?
```{r}
str(df)
```

-   Use the `summary()` function to get a summary of the dataframe. What insights can you gather from the summary?
```{r}
summary(df)
```

## Step 3: Accessing Data

-   Extract the `Sales` column as a vector using the `$` operator.
```{r}
df$Sales
```

-   Subset the first 15 rows and the columns `Order ID`, `Customer Name`, and `Sales`.
```{r}
df_subset <- df[1:15, c("Order.ID", "Customer.Name", "Sales")]
print(df_subset)
```


-   Use the `nrow()` and `ncol()` functions to determine the number of rows and columns in the dataframe.
```{r}
nrow(df)
ncol(df)
```

## Step 4: Filtering Data

-   Filter the dataframe to show only rows where the `Profit` is greater than 100.
```{r}
library(dplyr)
filtered_profit <- df %>% filter(Profit > 100)
print(filtered_profit)
```
```{r}
df[df$Profit > 100,]
```

-   Filter the dataframe to show only rows where the `Category` is `"Furniture"` and the `Sales` are greater than 500.
```{r}
filtered_furntiture_sales <- df %>% filter(Category == "Furniture" & Sales > 500)
print(filtered_furntiture_sales)
```
```{r}
df[df$Category == "Furniture" & df$Sales > 500, ]
```

-   Filter the dataframe to show only rows where the `Region` is `"West"` and the `Quantity` is greater than 5.
```{r}
filtered_west_quantity <- df %>% filter(Region == "West" & Quantity > 5)
print(filtered_west_quantity)
```
```{r}
df[df$Region == "West" & df$Quantity > 5, ]
```


## Step 5: Adding and Modifying Columns

-   Add a new column called `Profit Margin` that calculates the profit margin as `(Profit / Sales) * 100`.
```{r}
df <- transform(df, 
                'Profit.Margin' = (Profit/Sales) * 100
                )
head(df)
```

-   Modify the `Sales` column by rounding the values to 2 decimal places.
```{r}
df$Sales <- round(df$Sales, 2)
head(df)
```

-   Remove the `Postal Code` column from the dataframe using the `subset()` or `select()` function.

```{r}
df <- select(df, -Postal.Code)

```
```{r}
head(df)
```

## Step 6: Handling Missing Data

-   Check for missing values in the dataframe using the `is.na()` function. Are there any missing values?
```{r}
sum(is.na(df))
```
-   If there are missing values, remove rows with missing data using the `na.omit()` function.
```{r}
df <- na.omit(df)
```
-   Replace any missing values in the `Sales` column with the mean of the `Sales` column using the `na.fill()` function.
```{r}
library(zoo)
df$Sales <- na.fill(df$Sales, fill = mean(df$Sales))
```


## Step 7: Advanced Analysis (Optional)

-   Group the dataframe by `Region` and calculate the total `Sales` and `Profit` for each region.
```{r}
df_grouped <- df %>% 
  group_by(Region) %>%
  summarise(
    Total.Sales=sum(Sales),
    Total.Profit=sum(Profit)
  )
print(df_grouped)
```

-   Create a new column called `Discount Level` that categorizes the `Discount` column into:
    -   `"Low"` (0-0.2)
    -   `"Medium"` (0.2-0.5)
    -   `"High"` (0.5-1)
```{r}
df <- transform (df,
                 Discount.Level= case_when(
                   Discount >=0 & Discount <=0.2 ~ "Low",
                   Discount >=0.2 & Discount <=0.5 ~ "Medium",
                   Discount >=0.5 & Discount <=1 ~ "High",
                   TRUE ~ "Unknown" #In case of any other values outside the range
                 ))
head(df)
```

-   Sort the dataframe by `Sales` in descending order.
```{r}
df_sorted <- df %>%
  arrange(desc(Sales))
head(df_sorted)
```

